---
title: Exploring <Alice> and <Bob>’s Facebook Chat Messages within the Timeframe <start_date> - <end_date>
output:
  html_document:
    df_print: paged
  pdf_document: default
---

***

**Introduction**

***

**Libraries**

```{r libraries, warning=FALSE, message=FALSE}
library(plyr)
library(tidyverse)
library(rjson)
library(purrr)
library(lubridate)
library(data.table)
library(tidytext)
library(wordcloud)
library(RColorBrewer)
library(tm)
library(urltools)
```

```{r seed}
set.seed(2020)
```

***

**Source Files**

This version of the document assumes the message data was already acquired through the Facebook UI.

```{r messagepath, echo=FALSE}
# path to unzipped directory (named something like <BobSurname>_<RandomString>)
messages_path <- "your_message_path"
```

```{r namesetter}
sender1 <- "<Alice's Facebook screen name>"
sender2 <- "<Bob's Facebook screen name>"
```

```{r loader}
# concatenating the message files will take about a minute
filenames <- list.files(messages_path, pattern = "^message", full.names = TRUE)
messages <- filenames %>%
  map(~fromJSON(file = .)$messages) %>%
  unlist(recursive = FALSE) %>%
  map(data.frame) %>%
  rbind.fill()
messages$timestamp_ms <- as.POSIXct(messages$timestamp_ms / 1000, tz="CET", origin="1970-01-01")
messages$content <- as.character(messages$content)
```

***

**Message Count**

***Total count***
```{r mcount}
length(messages$content)
```

***Count per sender***
```{r msender}
messages %>%
  group_by(sender_name) %>%
  tally()
```

***Message count for each sender per month***
```{r monthandyear}
month_and_year <- function(timestamp) {
  substr(timestamp, 0, 7)
}
```

```{r bincount}
bins <- messages %>%
  group_by(month_and_year(timestamp_ms)) %>%
  n_groups()
```

```{r mlongitudinal}
ggplot(messages, aes(messages$timestamp_ms, fill=messages$sender_name)) +
  geom_histogram(position="stack", bins=bins) +
  labs(title = "Message Count per Month", x = "Month", y = "Message Count", fill = "Sender") +
  theme(legend.position = "bottom")
```

***Message count grouped by hour of the day***

```{r mtime}
ggplot(messages, aes(hour(messages$timestamp_ms), fill=messages$sender_name)) +
  geom_bar(position="dodge", width=0.8) +
  labs(title = "Message Count per Hour", x = "Hour", y = "Message Count", fill = "Sender") +
  theme(legend.position = "bottom")
```

***

**Message Length**

***Average length of messages per person***
```{r mlength}
messages$word_count <- messages$content %>%
  as.character() %>%
  strsplit(split = " ", fixed=TRUE) %>%
  map(length) %>%
  unlist

messages %>%
  group_by(sender_name) %>%
  summarise(avg_length = mean(word_count))
```

***Message length in time***
```{r mlengthlon}
word_count_time <- messages %>%
  group_by(month = month_and_year(timestamp_ms), sender_name) %>%
  summarise(avg_word_count = mean(word_count))
```

```{r plotmlengthlon}
ggplot(word_count_time, aes(x = month, y = avg_word_count, colour = sender_name)) +
  geom_line(group = word_count_time$sender_name) +
  labs(title = "Message Length per Month", x = "Month", y = "Average Message Length", colour = "Sender") +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1))
```

***

**Word Frequency Analysis**

The code in this section was largely adapted from the book [Text Mining with R](https://www.tidytextmining.com/tfidf.html).

```{r freq}
messages_words <- messages %>%
  unnest_tokens(word, content) %>%
  count(sender_name, word, sort = TRUE)

total_words <- messages_words %>%
  group_by(sender_name) %>%
  summarise(total = sum(n))

messages_words <- left_join(messages_words, total_words)
messages_words
```

```{r mtfidf}
messages_words <- messages_words %>%
  bind_tf_idf(word, sender_name, n)

messages_words %>%
  select(-total) %>%
  arrange(desc(tf_idf))
```

```{r mwordcloud}
plot_wordcloud <- function(word_df, max_words) {
  wordcloud(words = word_df$word, freq = word_df$tf_idf,
  max.words=max_words, random.order=FALSE, rot.per=0.35,
  colors=brewer.pal(8, "Dark2"))
}
```

***<Alice>-specific words***

```{r mwordcloud1}
# suppresses warnings about words that are excluded due to lack of space
suppressWarnings(messages_words %>%
  filter(sender_name == sender1) %>%
  drop_na(tf_idf) %>%
  plot_wordcloud(max_words = 50))
```

***<Bob>-specific words***

```{r mwordcloud2}
suppressWarnings(messages_words %>%
  filter(sender_name == sender2) %>%
  drop_na(tf_idf) %>%
  plot_wordcloud(max_words = 50))
```

It is likely necessary to perform more extensive data cleaning for all terms to make sense.

***

**Emoticon Usage**

This tracks emoticons like ":)" (rather than emoji). As many people have their own emoticon-related idiosyncrasies (remember "xD"?), this is a rather flaky code chunk.

The code in this section was largely adapted from the book [Text Mining with R](https://www.tidytextmining.com/tfidf.html).

```{r memoticon}
top_emoji <- messages %>%
  select(sender_name, content) %>%
  drop_na(content) %>%
  # do not tokenize words, as they remove useful punctuation
  separate_rows(content, sep = " ") %>%
  separate_rows(content, sep = "\n") %>%
  # filter out non-words
  filter(grepl("\\W", content)) %>%
  filter(!grepl("\\w[.,!?]", content)) %>%
  filter(!grepl("\\w['/:-]\\w", content)) %>%
  filter(!content %in% c(".", "?", "!", "...", "-", "*")) %>%
  count(sender_name, content, sort = TRUE) %>%
  group_by(sender_name) %>%
  top_n(10, n) %>%
  ungroup() %>%
  arrange(sender_name, -n) %>%
  mutate(content = reorder_within(content, n, sender_name))

ggplot(top_emoji, aes(content, n, fill = sender_name)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sender_name, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Most Common Emoticons", x = "Emoticon", y = "Count")
```

***Usage of "<3"***
```{r mhearts}
hearts <- messages %>%
  filter(grepl("<3", content, fixed = TRUE))

bincount <- round(interval(min(hearts$timestamp_ms), max(hearts$timestamp_ms)) / months(1))
ggplot(hearts, aes(hearts$timestamp_ms, fill=hearts$sender_name)) +
  geom_histogram(position="dodge", bins=bincount) +
  labs(title = "Heart Count per Month", x = "Month", y = "Heart Count", fill = "Sender") +
  theme(legend.position = "bottom")
```

***

**Shared Media**

***Shared links***

Number of shared links per person: 

```{r lsender}
messages$link <- domain(messages$link)

messages %>%
  drop_na(link) %>%
  group_by(sender_name) %>%
  tally()
```

Most commonly shared websites:

```{r lcount}
top_links <- messages %>%
  drop_na(link) %>%
  group_by(sender_name, link) %>%
  tally(sort = TRUE) %>%
  top_n(10, n) %>%
  ungroup() %>%
  arrange(sender_name, -n) %>%
  mutate(link = reorder_within(link, n, sender_name))

ggplot(top_links, aes(link, n, fill = sender_name)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sender_name, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  theme(axis.text.y = element_text(size=6)) +
  labs(title = "Commonly Shared Websites", x = "Website", y = "Count")
```

***Shared photos***

```{r psender}
n <- keep(names(messages), ~grepl("^photos.uri", .))

all_photos <- n %>%
  map(~select(messages, sender_name, timestamp_ms, .)) %>%
  map(drop_na) %>%
  rbindlist(use.names = FALSE)

all_photos %>%
  group_by(sender_name) %>%
  tally()
```

```{r pplotlon}
ggplot(all_photos, aes(timestamp_ms, fill=sender_name)) +
  geom_histogram(position="dodge", bins=bins) +
  labs(title = "Photo Count per Month", x = "Month", y = "Photo Count", fill = "Sender") +
  theme(legend.position = "bottom")
```

***Shared videos***

```{r vsender}
n <- keep(names(messages), ~grepl("^videos.uri", .))

n %>%
  map(~select(messages, sender_name, timestamp_ms, .)) %>%
  map(drop_na) %>%
  rbindlist(use.names = FALSE) %>%
  group_by(sender_name) %>%
  tally()
```

***Average call duration in seconds***

```{r cavg}
messages %>%
  filter(type == "Call") %>%
  filter(call_duration > 0) %>%
  summarise(avg_call_duration_s = mean(call_duration))
```

